Influence of Connectivity on Dung Beetle Communities

### 2025-06-14 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
writing extravanamza p3

Statistical analysis of abundance focused on the six most abundant species. A generalized linear mixed model identified significant effects for both species ID and patch type on dung beetle abundance (Table 7). The baseline abundance (intercept) corresponds to the abundance of _Aphodius alloblackburneus_ in connected patches. Compared to this baseline, _Canthon vigilans_, _Ateuchus lecontei_, and _Phanaeus igneus_ showed significantly higher abundance  (β = 1.05, 0.73, 0.66 respectively; all p < 0.001), while _Onthophagus pennsylvanicus_ had lower abundance (β = –0.80, p < 0.001).

Species-patch interactions indicated indicated species specific responses to treatments. In matrix plots, the positive effect on abundance was reduced for all other species (e.g., cvig × Matrix β = –0.59, p < 0.001). _Ateuchus lecontei_ and _Phanaeus igneus_ showed increased abundance in rectangle patches relative to expectations (β = 0.40 and 0.47, respectively; p < 0.05). Random effects indicated moderate variability among blocks (standard deviation of intercept = 0.51).

@@

The number of species per patch varied from N = 8 (rectangle patch in block 8) to N = 13 (matrix patch in block 53N). Species richness was relatively even across patch types with matrix and connected patches averaging out to approximately N = 11 species and rectangle and winged patches averaging to approximately N = 10 species (Figure 4). Overall there were no significant effects, however species richness patterns were variable between sampling blocks.

@@

Values for Hill numbers showed no patterns for shifts in species composition by patch type. Data for Shannon diversity showed no differences in the distribution of rare species between patch types. On average connected patches had a value of D = 5.29, while mean estimates for all other patches remained close with +0.03% for rectangle, -0.02% for matrix, and -0.06% for winged patches. There was no significance between patch types but data was highly variable between patch types and sampling blocks (Figure 5).

Values for Simpson Diversity also showed no differences but instead for more abundant species composing the dung beetle assemblages. Simpson diversity had a mean value of D = 4.07 in the connected patch, also with other values in close proximity with matrix +0.02%, rectangle +0.04%, and winged at -0.1%. There was no significant difference between patch types, however this data was also highly variable between patch type and sampling blocks (Figure 7).

### Functional Diversity

@@

>>>>

Statistical analysis of abundance focused on the six most abundant species. A generalized linear mixed model identified significant effects for both species ID and patch type on dung beetle abundance (Table 7). The baseline abundance (intercept) corresponds to the abundance of _Aphodius alloblackburneus_ in connected patches. Compared to this baseline, _Canthon vigilans_, _Ateuchus lecontei_, and _Phanaeus igneus_ showed significantly higher abundance  ($\beta$ = 1.05, 0.73, 0.66 respectively; all p < 0.001), while _Onthophagus pennsylvanicus_ had lower abundance ($\beta$ = –0.80, p < 0.001).

Species-patch interactions indicated indicated species specific responses to treatments. In matrix plots, the positive effect on abundance was reduced for all other species (e.g., cvig × Matrix $\beta$ = –0.59, p < 0.001). _Ateuchus lecontei_ and _Phanaeus igneus_ showed increased abundance in rectangle patches relative to expectations ($\beta$ = 0.40 and 0.47, respectively; p < 0.05). Random effects indicated moderate variability among blocks (standard deviation of intercept = 0.51).

@@

Plotting species richness by patch type reveals consistent richness across patch types with some variation between sampling blocks (Figure 5).The number of species per patch varied from N = 8 (rectangle patch in block 8) to N = 13 (matrix patch in block 53N). Modeling the effect of patch type on species richness with block as a random effect determined there was no significant differences among patch types. Comparing treatments using connected patches as a baseline resulted in no significant differences in matrix ($\beta$ = 0.05, p = 0.83), rectangle ($\beta$ = -0.07, p = 0.74), and winged patches ($\beta$ = -0.05, p = 0.83).

@@

We used linear mixed-effects models to compare the influence of patch type on both Shannon Diversity and Simpson's indexes, including block as a random effect to account for spatial variation. In both models, the reference level for patch type was Connected.

For Shannon Diversity, the estimated mean in Connected patches was 5.29 (SE = 0.66, t = 7.95). None of the alternative patch types showed statistically significant differences compared to Connected: Matrix ($\beta$ = –0.085, SE = 0.50, t = –0.17), Rectangle ($\beta$ = 0.15, SE = 0.50, t = 0.30), or Winged ($\beta$ = –0.34, SE = 0.50, t = –0.68).

Similarly, for Simpson's Diversity, the average value in Connected patches was 4.07 (SE = 0.64, t = 6.37). Again, none of the other patch types had significant effects: Matrix ($\beta$ = 0.11, SE = 0.41, t = 0.26), Rectangle ($\beta$ = 0.17, SE = 0.41, t = 0.40), and Winged ($\beta$ = –0.37, SE = 0.41, t = –0.91).

Across both models, the block-level random effect standard deviation was slightly greater than the residual error, indicating that variation between blocks accounted for a substantial portion of the overall variability.

@@

```{r Table M_rich, echo=FALSE,message = FALSE,warning=FALSE}

M_rich<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M_rich.csv") %>% 
  kable(
      digits = 3,
      align="lcc",
      format="latex",
      row.names = FALSE,
      booktabs=T,
      linesep = "", #removes the blank line after every 5 lines
      caption = "Richness Model") %>%
  kable_styling(bootstrap_options = c("striped"),
                full_width = F,
                latex_options=c("scale_down","hold_position"),
                font_size = 12,
                position = "center") %>%
  # column_spec(1, italic = T) %>%
  row_spec(0, bold = T)
  # row_spec(1:15, hline_after = T)
M_rich
```

```{r Table M_shan, echo=FALSE,message = FALSE,warning=FALSE}

M_shan<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M_shan.csv") %>% 
  kable(
      digits = 3,
      align="lcc",
      format="latex",
      row.names = FALSE,
      booktabs=T,
      linesep = "", #removes the blank line after every 5 lines
      caption = "Shannon Diversity Model") %>%
  kable_styling(bootstrap_options = c("striped"),
                full_width = F,
                latex_options=c("scale_down","hold_position"),
                font_size = 12,
                position = "center") %>%
  # column_spec(1, italic = T) %>%
  row_spec(0, bold = T)
  # row_spec(1:15, hline_after = T)
M_shan
```

```{r Table M_simp, echo=FALSE,message = FALSE,warning=FALSE}

M_simp<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M_simp.csv") %>% 
  kable(
      digits = 3,
      align="lcc",
      format="latex",
      row.names = FALSE,
      booktabs=T,
      linesep = "", #removes the blank line after every 5 lines
      caption = "Simpson's Index Model") %>%
  kable_styling(bootstrap_options = c("striped"),
                full_width = F,
                latex_options=c("scale_down","hold_position"),
                font_size = 12,
                position = "center") %>%
  # column_spec(1, italic = T) %>%
  row_spec(0, bold = T)
  # row_spec(1:15, hline_after = T)
M_simp
```
